\chapter*{Acknowledgements}

First of all, I would like to thank my supervisor, Juliano Wickboldt, for all the mentorship during the elaboration of this thesis. Also, for introducing me to the concept of the Digital Twin, which gave birth to the article. I also thank the Institute for parallel and distributed systems (IPVS) of the University of Stuttgart and the university itself for giving me the opportunity to learn many concepts I used in this work during an amazing research visit in Germany. Special thanks to my supervisors there, Pascal Hirmer and Ana Cristina Franco da Silva, who assisted me the whole year with mentorship and support. I also extend this acknowledgment to all the professors part of my student journey, who gave me the knowledge and instructions required for a Bachelor's degree.

Moreover, I thank my university colleagues, who I also proudly call my friends, for their support in the ups and downs of my university life. They were my agendas, coaches, free private tutors, role models, and psychologists many times. This includes all members of the (original) Whatsapp group CiC da Ousadia, but specially, Aline Weber, Catarina Nogueira, Guilherme Haetinger, and Marlize Ramos. 

Regarding my experiences outside the university, I kindly thank my instructors, leaders, colleagues, and friends of the Apple Developer Academy and do Sicredi. With them, I developed not just applications but also myself professionally and personally. 

Thank you to the friendships that remained from school and the exchange. Grateful to have shared many moments and feelings with them. Thanks especially to those who are or were in the same situation writing thesis, Diana Farje, Eduarda Mello, Lúcio Franco and Sofia Wender. Happy for this collegiality, even if unofficial, and for experiencing together (or almost that) this special moment that is graduation.

To conclude, special thanks to my brother, mother, and father. They were my safe harbor during my whole life and are always gonna be. Thank you for all the love, protection, support, and education it required to get here.

Thank you for standing with me when my code didn't work and celebrating with me when it did.

\chapter*{Agradecimentos}

Antes de tudo, gostaria de agradecer ao meu orientador, Juliano Wickboldt, por toda a mentoria durante a elaboração desta tese. Também, por me ter apresentado o conceito do Gêmeo Digital, que deu origem ao artigo. Agradeço também ao Instituto de Sistemas Paralelos e Distribuídos (IPVS) da Universidade de Stuttgart e à própria universidade por me terem dado a oportunidade de aprender muitos conceitos que utilizei neste trabalho durante um intercambio incrível na Alemanha. Agradecimentos especiais aos meus supervisores de lá, Pascal Hirmer e Ana Cristina Franco da Silva, que me acompanharam durante todo o ano. Também estendo este reconhecimento a todos os professores que participaram na minha jornada de estudante, que me forneceram os conhecimentos e instruções necessárias para a obtenção do diploma de Bacharelado.

Além disso, agradeço aos meus colegas universitários, a quem também chamo orgulhosamente meus amigos, pelo seu apoio nos altos e baixos da minha vida universitária. Foram muitas vezes as minhas agendas, coaches, professores particulares gratuitos, fontes de inspiração e psicólogos. Isto inclui todos os membros do grupo (original) Whatsapp CiC da Ousadia, mas especialmente, Aline Weber, Catarina Nogueira, Guilherme Haetinger, e Marlize Ramos. 

Quanto às minhas experiências fora da academia, agradeço gentilmente aos meus instrutores, líderes, colegas e amigos da Apple Developer Academy e Sicredi. Com eles, desenvolvi não só aplicativos, mas também a mim própria pessoa e profissionalmente. 

Obrigada às amizades que ficaram da escola e do intercâmbio. Grata por ter compartilhado muitos momentos e sentimentos com eles. Obrigada especialmente àqueles que estão ou estavam na mesma situação escrevendo tese, Diana Farje, Eduarda Mello, Lúcio Franco e Sofia Wender. Agradecida por este sentimento de coleguismo, mesmo que não oficial, e por podermos compartilhar juntos (ou quase isso) este momento especial que é a formatura.

Para concluir, agradecimentos especiais ao meu irmão, à minha mãe, e ao meu pai. Eles foram o meu porto seguro durante toda a minha vida e sempre o serão. Obrigada por todo o amor, protecção, apoio e educação necessários para chegar até aqui.

Obrigada a todos por estarem comigo nos momentos em que o meu código não compilou e por comemorarem comigo quando funcionou.

\clearpage